---
title: "Final_project6101"
author: "Jane"
date: "4/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load Package
```{r}
library(DT)
library(tidyverse)
library(ezids)
library(knitr)
library(kableExtra)
library(xtable)
library(Hmisc)
# The package "ezids" (EZ Intro to Data Science) includes a lot of the helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
library(ezids)
 
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times

```

```{r}
loadPkg("ggplot2")
```

# Featue Selection

From the EDA, features we've found out will be useful:

Categorical: 
Partner 
Tech.Support 
Online.Security 
Online.Backup 
Paperless.Billing 
Payment.Method 
Senior Citizen
Contract Type
Satisfaction Score
Dependents

Continuous:
Tenure
Avg. Monthly Charges

??? do we need have plots here as brief explanations as well as the rationale of the EDA and models??


# Model Selection

??? do we need to use model selection technique that we learned from class( like exhaustive test.etc) to determin model?
??? but  since the target variable has binary classifiers so Logistic reg

## 1. Logistic Reg

load data:
```{r}
tg <- data.frame(read.csv("~/GitHub/group7_project1/Data Files/Telco_customer_churn.csv")) 
head(tg)
xkablesummary(tg)

```

Categorical Variable

```{r}
tg$Churn.Value<- factor(tg$Churn.Value)
tg$Paperless.Billing <- factor(tg$Paperless.Billing)
tg$Payment.Method <- factor(tg$Payment.Method)
tg$Tech.Support <- factor(tg$Tech.Support)
tg$Online.Backup <- factor(tg$Online.Backup)
tg$Contract <- factor(tg$Contract)
tg$Internet.Service <- factor(tg$Internet.Service)
tg$Partner <- factor(tg$Partner)
tg$Dependents <- factor(tg$Dependents)
tg$Senior.Citizen <- factor(tg$Senior.Citizen)
```

```{r}

tgLogit <- glm(Churn.Value ~ Paperless.Billing+Payment.Method+Tech.Support+Online.Backup+Contract+Internet.Service+Partner+Senior.Citizen+Dependents , data = tg, family = "binomial")
```

```{r}
xkabledply(tgLogit, title = paste("Logistic Regression :", format(formula(tgLogit)) ))
```


```{r}
summary(tgLogit)
```



## Model Evaluation and comparsion
