---
title: "Group 7 Project 1: EDA"
author: "Andy Christian, Yaxin (Janet) Zhuang, Adhithya Kiran"
date: "3/28/2022"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
options(scientific=T, digits = 3)

#install packages
#install.packages("DT")
#install.packages("tidyverse")
#install.packages("ezids")
#install.packages("knitr")
#install.packages("kableExtra")
#install.packages("xtable")
#install.packages('Hmisc')

#load packages
library(DT)
library(tidyverse)
library(ezids)
library(knitr)
library(kableExtra)
library(xtable)
library(Hmisc)
```

# 1. Intro

*Loading Data Sets...*  
*Joining Data Sets...*  
*Selecting Variables...*  
*Converting Categorical Variables to Factor...*  
*Loading Functions...*  

```{r}
#Load data sets

#Main data set
churn_df <- data.frame(read_csv('Data Files/Telco_customer_churn.csv', col_types=cols()))

#additional data sets
churn_status_df <- data.frame(read_csv('Data Files/Telco_customer_churn_status.csv', col_types = cols()))
churn_services_df <- data.frame(read_csv('Data Files/Telco_customer_churn_services.csv', col_types = cols()))

```

```{r}
#Join and filter

#Filter columns to join from status_df
status_join <- churn_status_df %>%
  select(c(Customer.ID, Satisfaction.Score, Churn.Category))
#status_join

#Filer columns to join from services)df
service_join <- churn_services_df %>%
  select(c(Customer.ID, Avg.Monthly.Long.Distance.Charges, Avg.Monthly.GB.Download, Unlimited.Data, Total.Extra.Data.Charges))
#service_join

#Join status with churn
joined_churn_df <- merge(churn_df, status_join, by.x="CustomerID", by.y="Customer.ID")
#head(joined_churn_df)

#Join service with churn
joined_churn_df <- merge(joined_churn_df, service_join, by.x="CustomerID", by.y="Customer.ID")
#head(joined_churn_df)
```

```{r}
#drop unwanted columns from final df
filtered_joined_churn_df <- joined_churn_df %>%
  select(-c(1:10, 16, 20, 22:23, 28, 30:32))
```


```{r}
#Convert categorical to factor
cols <- c(1:3, 5:12, 14:17, 20)
filtered_joined_churn_df[cols] <- lapply(filtered_joined_churn_df[cols], as.factor)

# filtered_joined_churn_df$Senior.Citizen <- as.factor(filtered_joined_churn_df$Senior)#a
# filtered_joined_churn_df$Partner <- as.factor(filtered_joined_churn_df$Partner)
# filtered_joined_churn_df$Dependents <- as.factor(filtered_joined_churn_df$Dependents)#a
# filtered_joined_churn_df$Tech.Support <- as.factor(filtered_joined_churn_df$Tech)
# filtered_joined_churn_df$Internet.Service <- as.factor(filtered_joined_churn_df$Internet)#a
# filtered_joined_churn_df$Online.Security <- as.factor(filtered_joined_churn_df$Online.Security)
# filtered_joined_churn_df$Online.Backup <- as.factor(filtered_joined_churn_df$Online.Backup)
# filtered_joined_churn_df$Paperless.Billing <- as.factor(filtered_joined_churn_df$Paperless.Billing)
# filtered_joined_churn_df$Contract <- as.factor(filtered_joined_churn_df$Contract)#a
# filtered_joined_churn_df$Payment.Method <- as.factor(filtered_joined_churn_df$Payment.Method)
# filtered_joined_churn_df$Phone.Service <- as.factor(filtered_joined_churn_df$Phone.Service)
# filtered_joined_churn_df$Churn.Label <- as.factor(filtered_joined_churn_df$Churn.Label)
# filtered_joined_churn_df$Churn.Reason <- as.factor(filtered_joined_churn_df$Churn.Reason)
# filtered_joined_churn_df$Churn.Category <- as.factor(filtered_joined_churn_df$Churn.Category)
# filtered_joined_churn_df$Unlimited.Data <- as.factor(filtered_joined_churn_df$Unlimited.Data)

```

```{r}
#Seperate into churn and non-churn groupings for later comparison
only_churn <- joined_churn_df %>%
  filter(Churn.Value==1)

no_churn <- joined_churn_df %>%
  filter(Churn.Value==0)
```

```{r}
#Functions

#Create updated xkable_summary function to delete unwanted text
xkablesum_updated <- function (df, title = "Table: Statistics summary.", digits = 4,
          pos = "left", bso = "striped")
{
    s = summary(df)

    # Including RE NA's strip
    # Needed to add a dim check because including NA strip when no NA row made a new 7th row with text Min.:
    if (dim(s)[1] == 6) {
      strip_vector = c("Min.\\s*:\\s*", "1st Qu.\\s*:\\s*", "Median\\s*:\\s*", "Mean\\s*:\\s*",
                       "3rd Qu.\\s*:\\s*", "Max.\\s*:\\s*")
    }
    # NA's strip RE added here
    else if (dim(s)[1] == 7) {
      strip_vector = c("Min.\\s*:\\s*", "1st Qu.\\s*:\\s*", "Median\\s*:\\s*", "Mean\\s*:\\s*",
                       "3rd Qu.\\s*:\\s*", "Max.\\s*:\\s*", "NA's\\s*:\\s*")
    }

    # Made s = apply() -- without, didn't apply changes to table
    s <- apply(s, 2, function(x) stringr::str_remove_all(x, strip_vector))

    # Made s = apply()
    s <- apply(s, 2, function(x) stringr::str_trim(x, "right"))
    colnames(s) <- stringr::str_trim(colnames(s))

    if (dim(s)[1] == 6) {
        rownames(s) <- c("Min", "Q1", "Median",
                         "Mean", "Q3", "Max")
    }
    else if (dim(s)[1] == 7) {
        rownames(s) <- c("Min", "Q1", "Median",
                         "Mean", "Q3", "Max", "NA")
    }
    xkabledply(s, title = title, digits = digits, pos = pos,
               bso = bso)
}

#Better looking version of 2-sample t-test results than what the object itself displays
ttest2sample_info <- function(test) {
  if (test[["p.value"]] <= 0.05) {
  result = 'Reject the Null Hypothesis'
  } else {
    result = 'Do not reject the Null Hypothesis'
  }

  cat(c('\t', test$method, '\n\n',
      'Data:                       ', '|   ', test$data.name, '\n',
      'Null Hypothesis:            ', '|   true difference in means = ', test$null.value[1], '\n',
      'Alternative Hypothesis:     ', '|   true difference in means != ', test$null.value[1], '\n',
      'Confidence Level:           ', '|   ', attributes(test$conf.int)$conf.level, '\n',
      'Confidence Interval:        ', '|   [', round(test$conf.int[1], 2), ', ', round(test$conf.int[2], 2)), ']\n',
      'Sample Estimates of Mean:   ', '|   [X = ', test$estimate[1], ', Y = ', test$estimate[2], ']\n',
      'Test Values:                ', '|   [t = ', test$statistic,
                                   ', df = ', test$parameter[1],
                                   ', p-value = ', test[["p.value"]], ']\n',
      'Result:                     ', '|   ', result, sep='')
}

```

### 1.1 Data Set Overview

This data set is a fictional data set provided by IBM. It includes the demographic information on the customers of a telecommunications company, data on the services each customer used, charges paid (etc.), as well as whether the customer "churned" or not -- meaning, whether the customer *left the company in the past month*. It also includes data on the reason(s) each customer who did leave gave for leaving.

---

**From the graph below, we see that around 27% of customer left the platform within the last month.**

```{r}
options(repr.plot.width = 6, repr.plot.height = 4)
filtered_joined_churn_df %>%
group_by(Churn.Label) %>%
summarise(Count = n())%>%
mutate(percent = prop.table(Count)*100)%>%
ggplot(aes(reorder(Churn.Label, -percent), percent), fill = Churn)+
geom_col(fill = c("#FC4E07", "#E7B800"))+
geom_text(aes(label = sprintf("%.2f%%", percent)), hjust = 0.01,vjust = -0.5, size =3)+
theme_bw()+  
xlab("Churn") +
ylab("Percent")+
ggtitle("Churn Percent")
```

---

**The graph below details the reasons customers gave for leaving**.  
Although some customers may have selected the same *reason* for leaving, they may have *categorized* how they associated that reason differently. The churn category is used as a fill to provide further detail in understanding the customer perspective.  
We will use these reasons to help us formulate our questions, select appropriate variables to investigate, and contextualize the results:

```{r}
joined_churn_df %>%
  filter(!is.na(Churn.Category)) %>%
  ggplot() +
  geom_bar(mapping=aes(y=Churn.Reason, fill=Churn.Category), show.legend=TRUE) +
  labs(title='Reasons for Leaving by Answer', y=('Reason'), x=('Count')) +
  theme(plot.title = element_text(hjust = 0.5))

```

---

<br>

### 1.2 SMART Questions

Ultimately, the goal is to [produce a model for predicting customer churn]{.underline}. This initial project will focus on a basic EDA of the data set in preparation for building such a model. To Achieve this, three questions need to be aswered:

**1. *Which of the variables in the data set show a statistical difference between the churn and non-churn groupings?* **  
-This will require doing [2-sample t-tests]{.underline} for continuous variables, and [chi squared tests]{.underline} for categorical ones.  
-If a variable passes its test (i.e. p-value <= 0.05), that would suggest that there is a difference between the churn and non-churn groups with respect to that variable, and thus it likely has a measurable effect on the decision to churn, suggesting the variable could be used for modeling.  
--[This question will be answered in this project]{.underline}.

**2. *Of the variables that show a statistical difference between the churn and non-churn groupings, which are strong predictors of churn?* **  
-This will [require creating a model]{.underline}, for example, a logistical regression model, and examining the adjusted r^2 value.  
-[This question will be answered in Project 2]{.underline}.

**3. *Of the variables that are strong predictors of churn, which are actually providing useful information?* **  
-This will require critically thinking about [what the variable is actually telling us]{.underline}.  
-For example, it is likely that satisfaction score is more or less a corollary of churn and non-churn. Of course customers who churn are unsatisfied. Therefore, even though some variables may be strong predictors, they may not actually help us understand churn vs. non-churn.  
-[This question will be answered in Project 2]{.underline}.
<br>

### 1.3 Report Organization and Analysis Steps

1. Prepare and describe variables in data set  
2. Divide into churn and non-churn groupings
3. Examine descriptive statistics of the variables
4. Graph variables according to churn groupings to see if there are any obvious visual differences in the variables between churn groups  
5. Conduct the appropriate test for each variable to determine if there is a statistical difference between churn groups for each variable
6. Comment on the results of the analysis and propose next steps
<br>

# 2. Description of Data
<br>

### 2.1 Summary of Data Set Structure
<br>

#### **Initial Data Set**

Below is the structure of the data set before changing categorical variables to factors or dropping unwanted variables...

```{r results='markup', comment=NA}
#Note -- comment=NA removes the ## normally printed to output before each line

#display structure of unfiltered df -- so can discuss what was added and dropped
str(joined_churn_df)
#str(filtered_joined_churn_df)
```

#### **Selecting Variables**

The initial churn data frame `(Telco_customer_churn.csv)` (structure shown above) contained [`r ncol(churn_df)`]{.underline} variables. Another [`r ncol(service_join)-1`]{.underline} variables were added from the `Telco_customer_churn_services.csv` and  [`r ncol(status_join)-1`]{.underline} variables from the `Telco_customer_churn_status.csv` data set. Conversely, [18]{.underline} variables were dropped/not examined from the original `Telco_customer_churn.csv` data frame. The specific variables added and dropped, and the reason for doing so, follow below:
<br>

**Added Variables**

*From `Telco_customer_churn_status.csv`*

* ``r names(status_join)[2]``
* ``r names(status_join)[3]``

``r names(status_join)[2]`` was added because it matched reasons customer's gave for choosing to churn. Additionally the variable ``r names(status_join)[3]`` was added to allow for a more nuanced examination of the reasons customers gave for choosing to churn. Beyond selecting more variables to use in the investigation, knowing the specific reasons customers gave for leaving the company will be useful to compare against the results of the statistical analysis, in order to better understand and contextualize those results.

*From `Telco_customer_churn_services.csv`*

* ``r names(service_join)[2]``
* ``r names(service_join)[3]``
* ``r names(service_join)[4]``
* ``r names(service_join)[5]``

These variables were selected from the `Telco_customer_churn_services.csv` data set because they matched the reasons customers gave for choosing to churn, as recorded in the `Telco_customer_churn_status.csv` data set.

**Dropped/Not Examined Varaibles**

*From `Telco_customer_churn.csv`*

* ``r names(churn_df)[1]`` [not relevant]
* ``r names(churn_df)[2]`` [useful to ensure uniqueness, but otherwise not relevant]
* ``r names(churn_df)[3]`` [location data not examined]
* ``r names(churn_df)[4]`` [location data not examined]
* ``r names(churn_df)[5]`` [location data not examined]
* ``r names(churn_df)[6]`` [location data not examined]
* ``r names(churn_df)[7]`` [location data not examined]
* ``r names(churn_df)[8]`` [location data not examined]
* ``r names(churn_df)[9]`` [location data not examined]
* ``r names(churn_df)[10]`` [not relevant]
* ``r names(churn_df)[16]`` [assumed corollary with phone service, extra charge from more lines connected with monthly charges]
* ``r names(churn_df)[20]`` [didn't see a logical connection to the rest of the data]
* ``r names(churn_df)[22]`` [assumed corollary with GB download and internet speed (type)]
* ``r names(churn_df)[23]`` [assumed corollary with GB download and internet speed (type)]
* ``r names(churn_df)[28]`` [assumed corollary with monthly charges]
* ``r names(churn_df)[30]`` [Same as churn label]
* ``r names(churn_df)[31]`` [this is ultimately what we are trying to determine!]
* ``r names(churn_df)[32]`` [not examining profit, monthly charge more relevant for our purposes]

Most of the variables that were not examined were ignored because [they appeared obviously irrelevant, correlated with other variables, or did not seem to provide helpful information with respect to the questions we are trying to answer]{.underline}. Specifically regarding location data,  while examining location data could prove useful, it was decided to use it for follow up analysis. Part of the reason it was not chosen for the primary analysis is this data set does not contain data to dig further into what results on location analysis might mean. We don't have, for example, data on the strength of cell phone networks or availability of internet types by zip code, so it would be hard to usefully connect location data with other data and results.  
<br>

#### **Final Data Set**

The final data frame, `filtered_joined_churn_df`, consists of [`r ncol(filtered_joined_churn_df)`]{.underline} variables. Of these, [`r length(cols)`]{.underline} were *categorical* variables and converted into *factors*. There were [`r ncol(filtered_joined_churn_df) - length(cols)`]{.underline} *numerical* variables. Two of the variables, `Churn.Reason` and `Churn.Caegory` and included to help contextualize data and will not be analyzed. The variable `Churn.Label` is used to divide groups into churn and non-churn and will not itself be analyzed. Thus, there will be a total of [18]{.underline} variables that are analyzed in this EDA. The structure of the final data set is below:

```{r, comment=NA}
str(filtered_joined_churn_df)
```

**Brief Description of Final Set of Variables**

1.  ``r names(filtered_joined_churn_df)[1]`` [whether customer is a senior citizen]
2.  ``r names(filtered_joined_churn_df)[2]`` [whether customer is single]
3.  ``r names(filtered_joined_churn_df)[3]`` [whether customer has dependents]
4.  ``r names(filtered_joined_churn_df)[4]`` [length of time as customer]
5.  ``r names(filtered_joined_churn_df)[5]`` [whether customer using phone services]
6.  ``r names(filtered_joined_churn_df)[6]`` [type of internet service used -- affects speed]
7.  ``r names(filtered_joined_churn_df)[7]`` [whether customer uses online security service]
8.  ``r names(filtered_joined_churn_df)[8]`` [whether customer uses online backup service]
9.  ``r names(filtered_joined_churn_df)[9]`` [whether customer enrolled in tech support plan]
10. ``r names(filtered_joined_churn_df)[10]`` [type of contract (month-to-monthly, year, etc.)]
11. ``r names(filtered_joined_churn_df)[11]`` [whether customer uses paperless billing]
12. ``r names(filtered_joined_churn_df)[12]`` [payment method that customer uses]
13. ``r names(filtered_joined_churn_df)[13]`` [average total monthly charge for customer]
14. ``r names(filtered_joined_churn_df)[14]`` [whether a customer did or did not churn]
15. ``r names(filtered_joined_churn_df)[15]`` [specific reason customer gave for churning]
16. ``r names(filtered_joined_churn_df)[16]`` [rating customers gave for their satisfaction as a customer]
17. ``r names(filtered_joined_churn_df)[17]`` [for customers who churned, category of reason for leaving]
18. ``r names(filtered_joined_churn_df)[18]`` [average long-distance charges in a month]
19. ``r names(filtered_joined_churn_df)[19]`` [avg amount of data downloaded in a month]
20. ``r names(filtered_joined_churn_df)[20]`` [whether customer has unlimited data plan]
21. ``r names(filtered_joined_churn_df)[21]`` [lifetime extra data charges customer has accrued]
<br>

### 2.2 Display Joined and Filted Data Set
```{r results='markup'}
# filtered_joined_churn_df <- joined_churn_df %>%
#   select(-c(1:10, 16, 20, 22:23, 28, 30:33))

#display final filtered df  
DT::datatable(filtered_joined_churn_df,
         extensions = c('FixedColumns',"FixedHeader"),
          options = list(scrollX = TRUE,
                         paging=TRUE,
                         fixedHeader=TRUE))
#xkabledplyhead(joined_churn_df, title = "Churn Data Set")
```
<br>

### 2.3 Descriptive Statistics on Final Data Set
<br>

#### **Summary Tables of Categorical variables, Divided by Churn Groupings**
<br>

**All Customers**

```{r, comment=NA}
#Display summary statistics of categorical variables for entire df
filtered_joined_churn_df %>%
  select(c(cols, -Churn.Reason, -Churn.Category)) %>%
  summary()

```
<br>

**Churn Customers**

```{r, comment=NA}
#Display summary statistics of categorical variables for churn
filtered_joined_churn_df %>%
  filter(Churn.Label == "Yes") %>%
  select(c(cols, -Churn.Reason, -Churn.Category)) %>%
  summary()

```
<br>

**Non-Churn Customers**

```{r, comment=NA}
#Display summary statistics of categorical variables for non-churn
filtered_joined_churn_df %>%
  filter(Churn.Label == "No") %>%
  select(c(cols, -Churn.Reason, -Churn.Category)) %>%
  summary()

```

Focusing in on the Churn Customers vs. Non-Churn Customers summary tables, for many categorical variables [there appears to be a difference in the frequency of churn customers who belong to a given variable's level]{.underline}. For example, there are 1493 Churn Customers who are senior citizens and 476 who are not (~24%). For Non-Churn Customers, there are 4508 senior citizens and 666 non-seniors (~13%). While there are fewer non-seniors overall in the Churn Customer group, the percentage of seniors is higher compared to the Non-Churn group. The difference in frequency of each of these variable's levels will need to be analyzed.

<br>
<br>

#### **Summary Tables of Continuous variables, Divided by Churn Groupings**
<br>

**All Customers**
<br>

```{r results='markup', comment=NA}
#Display summary statistics of continuous variables for entire df
joined_churn_df %>%
  select(c(Tenure.Months, Monthly.Charges, Avg.Monthly.Long.Distance.Charges, Total.Extra.Data.Charges, Avg.Monthly.GB.Download)) %>%
  xkablesum_updated(title='Summary Statistics for Churn Data Set (ALL)')

```

**Churn Customers**
<br>

```{r results='markup', comment=NA}
#Compare with summary statistics of continuous variables for churn customers
joined_churn_df %>%
  filter(Churn.Value == 1) %>%
  select(c(Tenure.Months, Monthly.Charges, Avg.Monthly.Long.Distance.Charges, Total.Extra.Data.Charges, Avg.Monthly.GB.Download)) %>%
  xkablesum_updated(title='Summary Statistics for Churn Data Set (Churn/Left)')

```

**Non-Churn Customers**
<br>

```{r results='markup', comment=NA}
#Compare with summary statistics of continuous variables for non-churn customers
joined_churn_df %>%
  filter(Churn.Value == 0) %>%
  select(c(Tenure.Months, Monthly.Charges, Avg.Monthly.Long.Distance.Charges, Total.Extra.Data.Charges, Avg.Monthly.GB.Download)) %>%
  xkablesum_updated(title='Summary Statistics for Churn Data Set (Non-Churn/Stayed)')

```

At first glance, [there appears to be differences between the means of the Churn and Non-Churn groupings]{.underline} for `Tenure` and `Avg. Monthly Charges`. This will need to be analyzed further.
<br>
<br>

### 2.4 Another Useful Descriptive Statistics Function that Also Checks for NA's

From library `Hmisc`, using the `describe()` function

```{r, results='markup', comment=NA}

describe(filtered_joined_churn_df)

```

It should be noted that the only NA's that appear are in the `Churn.Label` and `Churn.Reason` variables. This is because only Churn Customers answered these questions. These NA values do not affect the analysis of the data set.

<br>
<br>

# 3. Graph Variables for Visual Comparison
<br>

### 3.1 Continuous Variables
<br>

---

#### **Tenure in Months**

```{r}
#Histograms
joined_churn_df %>%
  filter(Churn.Value==0) %>%
  ggplot() +
    geom_histogram(mapping=aes(x=Tenure.Months), fill='#CC6666', color='#9999CC', boundary=0, binwidth=1) +
    labs(title='Distribution of Length of Tenure (Non-Churn/Stayed)', x=('Number of Months'), y=('Count')) +
    theme(plot.title = element_text(hjust = 0.5))

joined_churn_df %>%
  filter(Churn.Value==1) %>%
  ggplot() +
    geom_histogram(mapping=aes(x=Tenure.Months), fill='#9999CC', color='#FFCCFF', boundary=0, binwidth=1) +
    labs(title='Distribution of Length of Tenure (Churn/Left)', x=('Number of Months'), y=('Count')) +
    theme(plot.title = element_text(hjust = 0.5))

#Boxplot
joined_churn_df %>%
  ggplot() +
    geom_boxplot(mapping=aes(y=Tenure.Months, fill=Churn.Label), color='#9999CC', boundary=0, binwidth=1) +
    labs(title='Distribution of Length of Tenure', x=('Tenure'), y=('Number of Months')) +
    theme(plot.title = element_text(hjust = 0.5,),
          axis.ticks.x = element_blank(),
          axis.text.x = element_blank())

#QQ plot non-churn
qqnorm(no_churn$Tenure.Months,
       main = "Tenure Q-Q Plot for Non-Churn Customers", ylab='Tenure Length Quantiles', col='#CC6666')
qqline(no_churn$Tenure.Months, col='#6633CC')

#QQ plot GRE
qqnorm(only_churn$Tenure.Months,
       main = "Tenure Q-Q Plot for Churn Customers", ylab='Tenure Length Quantiles', col='#9999CC')
qqline(only_churn$Tenure.Months, col='#663366')

```

**Visually, it looks like those who churned have a different tenure profile.**  
This can be seen in the shape of the histograms, as well as the fact that the mean lines in the boxplot of the churn and non-churn groups do not overlap with the other group's IQR.

Also -- the histogram and qq plot show a [clear lack of a normal distribution]{.underline}.

---

#### **Avg. Monthly Charges**

```{r}
#Histograms
joined_churn_df %>%
  filter(Churn.Value==0) %>%
  ggplot() +
    geom_histogram(mapping=aes(x=Monthly.Charges), fill='#CC6666', color='#9999CC', boundary=0, binwidth=5) +
    labs(title='Distribution of Avg. Monthly Charge (Non-Churn/Stayed)', x=('Monthly Charge'), y=('Count')) +
    theme(plot.title = element_text(hjust = 0.5))

joined_churn_df %>%
  filter(Churn.Value==1) %>%
  ggplot() +
    geom_histogram(mapping=aes(x=Monthly.Charges), fill='#9999CC', color='#FFCCFF', boundary=0, binwidth=5) +
    labs(title='Distribution of Avg. Monthly Charge (Churn/Left)', x=('Monthly Charge'), y=('Count')) +
    theme(plot.title = element_text(hjust = 0.5))

#Boxplot
joined_churn_df %>%
  ggplot() +
    geom_boxplot(mapping=aes(y=Monthly.Charges, fill=Churn.Label), color='#9999CC', boundary=0, binwidth=1) +
    labs(title='Distribution of Avg. Monthly Charge (All)', x=('Avg. Monthly Charge'), y=('Price')) +
    theme(plot.title = element_text(hjust = 0.5,),
          axis.ticks.x = element_blank(),
          axis.text.x = element_blank())

#QQ plot non-churn
qqnorm(no_churn$Monthly.Charges,
       main = "Avg. Monthly Charges Q-Q Plot for Non-Churn Customers", ylab='Avg. Monthly Charges Quantiles', col='#CC6666')
qqline(no_churn$Monthly.Charges, col='#6633CC')

#QQ plot GRE
qqnorm(only_churn$Monthly.Charges,
       main = "Avg. Monthly Charges Q-Q Plot for Churn Customers", ylab='Avg. Monthly Charges Quantiles', col='#9999CC')
qqline(only_churn$Monthly.Charges, col='#663366')
```

**Visually, it looks like those who churned have a different price profile.**  
This can be seen in the shape of the histograms, and while the mean lines in the boxplot of the churn and non-churn groups do overlap with the other group's IQR, they still appear to be quite different.

Also -- the histogram and qq plot show a [clear lack of a normal distribution]{.underline}.

---

#### **Avg. Monthly Long-Distance Charges**

```{r}
#Histograms
joined_churn_df %>%
  filter(Churn.Value==0) %>%
  ggplot() +
    geom_histogram(mapping=aes(x=Avg.Monthly.Long.Distance.Charges), fill='#CC6666', color='#9999CC', boundary=0, binwidth=5) +
    labs(title='Distribution of Avg. Monthly Long-Distance Charge (Non-Churn/Stayed)', x=('Monthly Charge'), y=('Count')) +
    theme(plot.title = element_text(hjust = 0.5))

joined_churn_df %>%
  filter(Churn.Value==1) %>%
  ggplot() +
    geom_histogram(mapping=aes(x=Avg.Monthly.Long.Distance.Charges), fill='#9999CC', color='#FFCCFF', boundary=0, binwidth=5) +
    labs(title='Distribution of Avg. Monthly Long-Distance Charge (Churn/Left)', x=('Monthly Charge'), y=('Count')) +
    theme(plot.title = element_text(hjust = 0.5))

#Boxplot
joined_churn_df %>%
  ggplot() +
    geom_boxplot(mapping=aes(y=Avg.Monthly.Long.Distance.Charges, fill=Churn.Label), color='#9999CC', boundary=0, binwidth=1) +
    labs(title='Distribution of Avg. Monthly Long-Distance Charge (All)', x=('Avg. Monthly Long-Distance Charge'), y=('Price')) +
    theme(plot.title = element_text(hjust = 0.5,),
          axis.ticks.x = element_blank(),
          axis.text.x = element_blank())

#QQ plot non-churn
qqnorm(no_churn$Avg.Monthly.Long.Distance.Charges,
       main = "Avg. Monthly Long-Distance Charge Q-Q Plot for Non-Churn Customers", ylab='Avg. Monthly Long-Distance Charge Quantiles', col='#CC6666')
qqline(no_churn$Avg.Monthly.Long.Distance.Charges, col='#6633CC')

#QQ plot GRE
qqnorm(only_churn$Avg.Monthly.Long.Distance.Charges,
       main = "Avg. Monthly Long-Distance Charge Q-Q Plot for Churn Customers", ylab='Avg. Monthly Long-Distance Charge Quantiles', col='#9999CC')
qqline(only_churn$Avg.Monthly.Long.Distance.Charges, col='#663366')
```

**Visually, it does not look like those who churned have a difference in their avg. monthly long distance charges.**  
The shape of the churn and non-churn histograms for long distance charges look similar. Their boxplots are also nearly identical.

Also -- the histogram and qq plot show a [clear lack of a normal distribution]{.underline}.

---

#### **Avg. Monthly GB Download**

```{r}
#Histograms
joined_churn_df %>%
  filter(Churn.Value==0) %>%
  ggplot() +
    geom_histogram(mapping=aes(x=Avg.Monthly.GB.Download), fill='#CC6666', color='#9999CC', boundary=0, binwidth=5) +
    labs(title='Distribution of Avg Monthly GB Download (Non-Churn/Stayed)', x=('Avg Monthly GB Download'), y=('Count')) +
    theme(plot.title = element_text(hjust = 0.5))

joined_churn_df %>%
  filter(Churn.Value==1) %>%
  ggplot() +
    geom_histogram(mapping=aes(x=Avg.Monthly.GB.Download), fill='#9999CC', color='#FFCCFF', boundary=0, binwidth=5) +
    labs(title='Distribution of Avg. Monthly GB Download (Churn/Left)', x=('Avg Monthly GB Download'), y=('Count')) +
    theme(plot.title = element_text(hjust = 0.5))

#QQ plot non-churn
qqnorm(no_churn$Avg.Monthly.GB.Download,
       main = "Avg. Monthly GB Download Q-Q Plot for Non-Churn Customers", ylab='Avg. Monthly GB Download Quantiles', col='#CC6666')
qqline(no_churn$Monthly.Charges, col='#6633CC')

#QQ plot GRE
qqnorm(only_churn$Avg.Monthly.GB.Download,
       main = "Avg. Monthly GB Download Q-Q Plot for Churn Customers", ylab='Avg. Monthly GB Download Quantiles',col='#9999CC')
qqline(only_churn$Avg.Monthly.GB.Download, col='#663366')
```

**Visually, it looks like those who left have a different Avg monthly gb downloaded profile.**  
This can be seen in the shape of the histograms.

Also -- the histogram and qq plot show a [clear lack of a normal distribution]{.underline}.

---

#### **Total Extra Data Charges**

```{r}
#Histograms
joined_churn_df %>%
  filter(Churn.Value==0) %>%
  ggplot() +
    geom_histogram(mapping=aes(x=Total.Extra.Data.Charges), fill='#CC6666', color='#9999CC', boundary=0, binwidth=5) +
    labs(title='Distribution of Total Extra Data Charges (Non-Churn/Stayed)', x=('Total Extra Data Charges'), y=('Count')) +
    theme(plot.title = element_text(hjust = 0.5))

joined_churn_df %>%
  filter(Churn.Value==1) %>%
  ggplot() +
    geom_histogram(mapping=aes(x=Total.Extra.Data.Charges), fill='#9999CC', color='#FFCCFF', boundary=0, binwidth=5) +
    labs(title='Distribution of Total Extra Data Charges (Churn/Left)', x=('Total Extra Data Charges'), y=('Count')) +
    theme(plot.title = element_text(hjust = 0.5))

#QQ plot non-churn
qqnorm(no_churn$Total.Extra.Data.Charges,
       main = "Total Extra Data Charges Q-Q Plot for Non-Churn Customers", ylab='Total Extra Data Charges Quantiles', col='#CC6666')
qqline(no_churn$Monthly.Charges, col='#6633CC')

#QQ plot GRE
qqnorm(only_churn$Total.Extra.Data.Charges,
       main = "Avg. Total Extra Data Charges Q-Q Plot for Churn Customers", ylab='Total Extra Data Charges', col='#9999CC')
qqline(only_churn$Total.Extra.Data.Charges, col='#663366')
```

**Visually, it looks like those who left do not have a different Total Extra Data Charges profile.**  
This can be seen in the shape of the histograms.

Also -- the histogram and qq plot show a [clear lack of a normal distribution]{.underline}.

---

<br>

### 3.2 Categorical Variables
<br>

---

#### **Senior Citizen**

```{r}
joined_churn_df %>%
  ggplot() +
  geom_bar(mapping=aes(x=Churn.Label, fill=Senior.Citizen), position = 'fill') +
  labs(title='Frequency of Being Senior Citizen in Churn Groupings', x=('Churn or Non-Churn'), y=('Per of Group is Senior')) +
  theme(plot.title = element_text(hjust = 0.5))

```

**Visually, there appears to be more senior citizens in the churn group than the non-churn group.**  
This suggests being a senior affected whether someone decided to churn.

---

#### **Partner**

```{r}
options(repr.plot.width = 12, repr.plot.height = 8)
ggplot(joined_churn_df, aes(x=Partner,fill=Churn.Label))+ geom_bar(position = 'fill')
```

**This plot shows customers with Partners have lower churn rate as compared to those who don't have partners.And This indicated that whether have partner or not could affect customers make decision on whether to churn or not churn**

---

#### **Dependents**

```{r}
options(repr.plot.width = 12, repr.plot.height = 8)
ggplot(joined_churn_df, aes(x=Dependents,fill=Churn.Label))+ geom_bar(position = 'fill')
```


**Visually, there appears to be more non dependents in the churn group than the non-churn group.**  
This suggests being a dependency affected whether someone decided to churn.

---

#### **Phone Service**

```{r}
options(repr.plot.width = 12, repr.plot.height = 8)
ggplot(joined_churn_df, aes(x=Phone.Service,fill=Churn.Label))+ geom_bar(position = 'fill')
```

**Visually, there appears to be almost equal profiles having phone service and not having phone service in the churn group and in the non-churn group.**  
This suggests being having a phone service not affected whether someone decided to churn.

---

#### **Internet Service**

```{r}
options(repr.plot.width = 12, repr.plot.height = 8)
ggplot(joined_churn_df, aes(x=Internet.Service,fill=Churn.Label))+ geom_bar(position = 'fill')
```


**Visually, there appears to be more profiles having fiber optics in the churn group than the non-churn group.**  
This suggests being a Internet services affected whether someone decided to churn.

---

#### **Online Security**
```{r}
ggplot(joined_churn_df, aes(x=Online.Security,fill=Churn.Label))+ geom_bar(position = 'fill')+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10))
```

**This plot represents customers who do not have services like Online Security have left the platfom as conpared to other options.And This indicated that the option of Online Security could affect customers make decision on whether to churn or not churn**

---

#### **Online Backup**
```{r}
 ggplot(joined_churn_df, aes(x=Online.Backup,fill=Churn.Label))+ geom_bar(position = 'fill')+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10))
```

**The graph shows customers who do not have services like Online Backup have left the platform. And This indicated that the option of Online Backup could affect customers make decision on whether to churn or not churn**

---

#### **Tech Support**
```{r}
ggplot(joined_churn_df, aes(x=Tech.Support,fill=Churn.Label))+ geom_bar(position = 'fill')+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10))
```

**The plot shows customers who do not have services like Tech Support have left the platform in the past month.And This indicated that the option of Tech Support could affect customers make decision on whether to churn or not churn**

---

#### **Contract Type**

```{r}
joined_churn_df %>%
  ggplot() +
  geom_bar(mapping=aes(x=Churn.Label, fill=Contract), position = 'fill') +
  labs(title='Frequency of Contract Type in Churn Groupings', x=('Churn or Non-Churn'), y=('Per of Group has Contract Type')) +
  theme(plot.title = element_text(hjust = 0.5))

```

**Visually, there appears to be far more month-to-month contract customers in the churn group than the non-churn group.**  

This suggests contract type affected whether someone decided to churn.

---

#### **Paperless Billing**
```{r}
 ggplot(joined_churn_df, aes(x=Paperless.Billing,fill=Churn.Label))+
          geom_bar(position = 'fill')+theme_bw()+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10))
```

**The plot represents that churn percent is higher in case of customers having paperless billing option. And This indicated that the option of Paperless Billing could affect customers make decision on whether to churn or not churn**

---

#### **Payment Method**
```{r}
ggplot(joined_churn_df, aes(x=Payment.Method,fill=Churn.Label))+
          geom_bar(position = 'fill')+theme_bw()+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10))
```

**The plot represents that customers who have Electronic Check Payment Method tend to leave the platform more when compared to other options. And This indicated that the option of payment method could affect customers make decision on whether to churn or not churn**

---

#### **Satisfaction Score**

```{r}
#Had to add it as.factor()
joined_churn_df %>%
  ggplot() +
  geom_bar(mapping=aes(x=Churn.Label, fill=as.factor(Satisfaction.Score)), position = 'fill') +
  labs(title='Frequency Satisfaction Rating in Churn Groupings', x=('Churn or Non-Churn'), y=('Per of Group Gave Satisfaction Rating')) +
  theme(plot.title = element_text(hjust = 0.5))

```

**Visually, there is little overlap in the ratings given by the churn group compared to the non-churn group.**  
This suggests that the rating given by customers affects whether someone decided to churn.  

---

#### **Unlimited Data**

```{r}
#Had to add it as.factor()
joined_churn_df %>%
  ggplot() +
  geom_bar(mapping=aes(x=Churn.Label, fill=as.factor(Unlimited.Data)), position = 'fill') +
  labs(title='Frequency Unlimited Data in Churn Groupings', x=('Churn or Non-Churn'), y=('Per of Group has Unlimited Data')) +
  theme(plot.title = element_text(hjust = 0.5))

```

**Visually, those who have unlimited data churned morethan those who have not **  

This suggests the frequency of unlimited data affects whether someone decided to churn.

---

<br>

# 4. Tests for Variables to Check for Statistical Signficance of Differences/Independence
<br>

### 4.1 Continuous Variables -- 2-sample T-tests

For each continuous variable being examined, the means of the churn and the non-churn groupings are being compared. Thus, [2-sample t-tests]{.underline} are used to see if the actual difference of the means between the churn groups is 0. So, for all variables tested, the hypothesis test is:  

**NULL HYPOTHESIS:**  
The true difference in the means *between the `churn` and the `non-churn` group* [is 0]{.underline}.

A *p-value* [<= 0.05]{.underline} will result in this null hypothesis being [rejected]{.underline} and there [being sufficient evidence]{.underline} to argue that there is a *real difference in the means* of the churn and non-churn groups for a given variable.
<br>

---

#### **Tenure in Months**

```{r results='markdown', comment=NA}
ttest2sample_tenure = t.test(only_churn$Tenure.Months,no_churn$Tenure.Months)
#cat("\nt-test object's display of results\n")
#ttest2sample_tenure
#cat("\nAndy's function's display of results\n\n")
ttest2sample_info(ttest2sample_tenure)
```

**The results of the test returned a [p-value less than 0.05]{.underline}. As a result, we [reject the Null Hypothesis]{.underline}.**  
Since there is a real difference in the means of the churn and non-churn groups with respect to their `Tenure`, this suggests `Tenure` affects whether a customer chooses to churn or not.  

---

#### **Avg. Monthly Charges**

```{r results='markup', comment=NA}
ttest2sample_monthlyCharge = t.test(only_churn$Monthly.Charges,no_churn$Monthly.Charges)
#ttest2sample_monthlyCharge
ttest2sample_info(ttest2sample_monthlyCharge)

```

**The results of the test returned a [p-value less than 0.05]{.underline}. As a result, we [reject the Null Hypothesis]{.underline}.**  
Since there is a real difference in the means of the churn and non-churn groups with respect to their `Avg. Monthly Charges`, this suggests `Avg. Monthly Charges` affects whether a customer chooses to churn or not.  

---

#### **Avg. Monthly Long-Distance Charges**

```{r results='markup', comment=NA}
ttest2sample_longDistance = t.test(only_churn$Avg.Monthly.Long.Distance.Charges,no_churn$Avg.Monthly.Long.Distance.Charges)
#ttest2sample_longDistance
ttest2sample_info(ttest2sample_longDistance)

```

**The results of the test returned a [p-value greater than 0.05]{.underline}. As a result, we [accept the Null Hypothesis]{.underline}.**  
Since we cannot say that there is a real difference in the means of the churn and non-churn groups with respect to their `Avg. Monthly Long Distance Charges`, this suggests `Avg. Monthly Long Distance Charges` does not affect whether a customer chooses to churn or not.  

---

#### **Avg. Monthly GB Download**

```{r results='markup', comment=NA}
ttest2sample_AMG = t.test(only_churn$Avg.Monthly.GB.Download ,no_churn$Avg.Monthly.GB.Download )
#ttest2sample_AMG
ttest2sample_info(ttest2sample_AMG)

```

**The results of the test returned a [p-value less than 0.05]{.underline}. As a result, we [reject the Null Hypothesis]{.underline}.**  
Since there is a real difference in the means of the churn and non-churn groups with respect to their `Avg.Monthly.GB.Download`, this suggests `Avg.Monthly.GB.Download` affects whether a customer chooses to churn or not.

---

#### **Total Extra Data Charges**


```{r results='markup', comment=NA}
ttest2sample_Ex = t.test(only_churn$Total.Extra.Data.Charges ,no_churn$Total.Extra.Data.Charges)
#ttest2sample_Ex
ttest2sample_info(ttest2sample_Ex)

```

**The results of the test returned a [p-value greater than 0.05]{.underline}. As a result, we [accept the Null Hypothesis]{.underline}.**  
Since there is no real difference in the means of the churn and non-churn groups with respect to their `Total.Extra.Data.Charges`, this suggests `Total.Extra.Data.Charges` do not affects whether a customer chooses to churn or not.

---

<br>

### 4.2 Categorical Variables -- Goodness of Fit and Test of Independence
<br>

---

#### **Senior Citizen**

Since `Senior.Citizen` only has two levels, a chi squared, [goodness of fit]{.underline} test was used. A numerical vector of the count of senior/non-senior customers in the Churn Group was passed as the x parameter. The frequency of senior/non-senior customers in the Non-Churn Group was used as the 'expected' frequency and passed as the p parameter. Thus the hypothesis is as follows:  

**NULL HYPOTHESIS:**  
There is no difference in frequency of senior/non-senior between the Churn and Non-Churn Groups.

<br>

*Numerical Vector, Senior/Non-Senior in [Churn Group]{.underline}*
```{r results='markup', comment=NA}
#Numerical vector -- Count of senior, non-senior in the Churn group
count_churn_senior <- only_churn %>%
  group_by(Senior.Citizen) %>%
  summarise(count = n())
count_churn_senior
```
<br>

*Freq Distribution of Senior/Non-Senior in [Non-Churn Group]{.underline} (the 'expected frequency` that churn is being compared against)*
```{r results='markdown', comment=NA}
#Count of senior, non-senior in the non-churn group
freq_nonchurn_senior <- no_churn %>%
  group_by(Senior.Citizen) %>%
  summarise(count = n())

#Find 'expected' frequency -- i.e. freq in non-churn group
perc_freq_senior = freq_nonchurn_senior$count/nrow(no_churn)
perc_freq_senior

```
<br>

*Results of Chi Squared, Goodness of Fit Test*
```{r results='markdown', comment=NA}
#Use chi sqr test to compare 'expected' freq (freq of seniors for non-churn) with freq of seniors for churn
#low p-value means their frequencies are different.
chigoodness_senior = chisq.test(count_churn_senior$count, p=perc_freq_senior)
chigoodness_senior

```

**The results of the test returned a [p-value less than 0.05]{.underline}. As a result, we [reject the Null Hypothesis]{.underline}.**  
Since there is a difference in the frequency of `Senior Citizen` between the churn and non-churn groups, this suggests that whether a customer is a `Senior Citizen` affects whether they choose to churn or not.

---

#### **Partner**
```{r, comment=NA}
test_partner <- chisq.test(table(joined_churn_df$Churn.Label, joined_churn_df$Partner))
test_partner
```

**p-value <2e-16, it is not independent. The two variables relate to each other this indicate that whether have partner or not relate to customers make decision on whether to churn or not churn**

---

#### **Dependents**

```{r, comment=NA}
test_Dependents <- chisq.test(table(joined_churn_df$Churn.Label, joined_churn_df$Dependents))
test_Dependents
```

**p-value <2e-16, it is not independent. The two variables relate to each other**

---

#### **Phone Service**

```{r, comment=NA}
test_PS <- chisq.test(table(joined_churn_df$Churn.Label, joined_churn_df$Phone.Service))
test_PS
```

**p-value = .03, it is independent. The two variables did not relate to each other**

---



```{r, comment=NA}
#test_IS <- chisq.test(table(joined_churn_df$Churn.Label, joined_churn_df$Internet.Service))
#test_IS
```


---

#### **Online Security**
```{r ,results='markup'}
x1 <- table(joined_churn_df$Churn.Label,joined_churn_df$Online.Security)
```

```{r,results='hold', comment=NA}
chitest_os <- chisq.test(x1)
chitest_os

#p-value <2e-16, it is not independent. The two variables relate to each other.
```

**p-value <2e-16, it is not independent. The two variables relate to each other this indicate that the option of online security relate to customers make decision on whether to churn or not churn**

---

#### **Online Backup**
```{r ,results='markup'}
x2 <- table(joined_churn_df$Churn.Label,joined_churn_df$Online.Backup)
```

```{r,results='hold', comment=NA}
chitest_ob <- chisq.test(x2)
chitest_ob

#p-value <2e-16, it is not independent. The two variables relate to each other.
```

**p-value <2e-16, it is not independent. The two variables relate to each other this indicate that the option of online backup relate to customers make decision on whether to churn or not churn**

---

#### **Tech Support**
```{r ,results='markup'}
x3 <- table(joined_churn_df$Churn.Label,joined_churn_df$Tech.Support)
```

```{r,results='hold', comment=NA}
chitest_ts <- chisq.test(x3)
chitest_ts
#p-value <2e-16, it is not independent. The two variables relate to each other.
```

**p-value <2e-16, it is not independent. The two variables relate to each other this indicate that the option of tech support relate to customers make decision on whether to churn or not churn**

---

#### **Contract Type**

Since `Contract.Type` has more than two levels, a chi squared, [test of independence]{.underline} was used. A contingency table was created comparing Churn/Non-Churn against the levels of Contract. Then, this table was passed to the chisq.test() function to get the result of the test. Thus the hypothesis is as follows:  

**NULL HYPOTHESIS:**  
Contract Type is independent of Churn.

```{r results='markup'}
#Create contingency table
contable_contract = table(as.integer(joined_churn_df$Churn.Value), filtered_joined_churn_df$Contract)
xkabledply(contable_contract, title="Contingency table for Churn Against Contract Type")

```

*Results*
```{r, comment=NA}
#conduct independence test with contingency table
chitest_contract = chisq.test(contable_contract)
chitest_contract

```

**The results of the test returned a [p-value less than 0.05]{.underline}. As a result, we [reject the Null Hypothesis]{.underline}.**  
`Churn` and `Contract Type` are not independent of one another. Thus, this suggests that the type of `Contract` a customer has affects whether they choose to churn or not.

---

#### **Paperless Billing**
```{r, comment=NA}
test2 <- chisq.test(table(joined_churn_df$Churn.Label, joined_churn_df$Paperless.Billing))
test2
#p-value <2e-16, it is not independent. The two variables relate to each other.

```

**p-value <2e-16, it is not independent. The two variables relate to each other  this indicate that the option of paperless billing relate to customers make decision on whether to churn or not churn**

---

#### **Payment Method**
```{r ,results='markup'}
x4 <- table(joined_churn_df$Churn.Label,joined_churn_df$Payment.Method)
```

```{r,results='hold', comment=NA}
chitest_pm <- chisq.test(x4)
chitest_pm

#p-value <2e-16, it is not independent. The two variables relate to each other.
```

**p-value <2e-16, it is not independent. The two variables relate to each other this indicate that the option of payment method relate to customers make decision on whether to churn or not churn**

---

#### **Unlimited Data**

```{r ,results='markup'}
x5 <- table(joined_churn_df$Churn.Label,joined_churn_df$Unlimited.Data)


```

```{r,results='hold'}
chitest_pm <- chisq.test(x5)
chitest_pm

#p-value <2e-16, it is not independent. The two variables relate to each other.



```

**p-value <2e-16, it is not independent. The two variables relate to each other**

---


#### **Satisfaction Score**

Since `Satisfaction Score` has more than two levels, a chi squared, [test of independence]{.underline} was used. A contingency table was created comparing Churn/Non-Churn against the Satisfaction Ratings. Then, this table was passed to the chisq.test() function to get the result of the test. Thus the hypothesis is as follows:  

**NULL HYPOTHESIS:**  
Satisfaction Score is independent of Churn.

```{r results='markup'}
#Create contingency table
contable_sat = table(as.integer(joined_churn_df$Churn.Value), filtered_joined_churn_df$Satisfaction.Score)
xkabledply(contable_sat, title="Contingency table for Churn Against Satisfaction Score")

```

*Results*
```{r, comment=NA}
#conduct independence test with contingency table
chitest_sat = chisq.test(contable_sat)
chitest_sat

```

**The results of the test returned a [p-value less than 0.05]{.underline}. As a result, we [reject the Null Hypothesis]{.underline}.**  
`Churn` and `Satisfaction Score` are not independent of one another. Thus, this suggests that a customer's level of `Satisfaction` affects whether they choose to churn or not.

---

#### **Internet Services**

Since `Internet Services` has more than two levels, a chi squared, [test of independence]{.underline} was used. A contingency table was created comparing Churn/Non-Churn against the different types of Internet Services. Then, this table was passed to the chisq.test() function to get the result of the test. Thus the hypothesis is as follows:  

**NULL HYPOTHESIS:**  
Internet Services is independent of Churn.

```{r results='markup'}
#Create contingency table
contable_IS = table(as.integer(joined_churn_df$Churn.Value), filtered_joined_churn_df$Internet.Service)
xkabledply(contable_IS, title="Contingency table for Churn Against Internet Service")

```

*Results*
```{r, comment=NA}
#conduct independence test with contingency table
chitest_IS = chisq.test(contable_sat)
chitest_IS

```

**The results of the test returned a [p-value less than 0.05]{.underline}. As a result, we [reject the Null Hypothesis]{.underline}.**  
`Churn` and `Internet Service` are not independent of one another. Thus, this suggests that a customer's level of `Internet Service` affects whether they choose to churn or not.

---

# 5. Conclusion

<br>

### 5.1 Answering SMART Question 1

**1. *Which of the variables in the data set show a statistical difference between the churn and non-churn groupings?* **

To answer this question, we divided the data into churn and non-churn groups, then visually explored how the data set's variables differed when divided into these groups. We then tested the differences using 2-sample t-tests, goodness of fit chi squared tests, and chi squared tests of independence. From this we determined that the following [15]{.inderline} variables *showed a statistical difference between the churn and non-churn groups:*

* Tenure
* Avg. Monthly Charges
* Avg. Monthly GB Download
* Senior Citizen
* Partner
* Dependents
* Internet Service
* Online Security
* Online Backup
* Tech Support
* Contract Type
* Paperless Billing
* Payment Method
* Satisfaction Score
* Unlimited Data

### 5.2 Looking Foward to Model Building and Ansering SMART Questions 2 & 3

As we move into Project 2, we will need to critically examine the 15 variables that passed our tests to see if they help us build a model to predict the likelihood of a customer churning. We will be answering the following questions:

---

**2. *Of the variables that show a statistical difference between the churn and non-churn groupings, which are strong predictors of churn?* **

**3. *Of the variables that are strong predictors of churn, which are actually providing useful information?* **  

---

While these variables showed a statistical difference between the churn and non-churn groups, that does not necessarily mean that they are good for model building or tell us useful information. As we move forward with our project, we will need to contextualize the analysis we have done here in order to select variables for our model that actually make sense in helping us answer our question about predicting whether a customer will churn. It is possible that some variables, while looking good for a model, don't logically connect to our question and may actually mask those variables more important to finding our answer. This will all be explored, evaluated, and addressed in our second project.
<br>

For example, from the earlier analysis on Techsupport, online backup and online security. Customers who do not have service like online backup, online security and tech support have higher churn rate as compared to other options. This make a intuitive sense that customers are not willing to stay with a telecom platform that have poor performance on customer support system. Therefore, I assume these three variables have some capability to predict some part of target value. 
<br>

# 6. References

Chang, Jack. (2020). *Telco customer churn (11.1.3+)* [Data Set]. Kaggle. https://www.kaggle.com/datasets/ylchang/telco-customer-churn-1113

Samples Team. (2019, July 11). *Telco customer churn (11.1.3+)*. IBM. https://community.ibm.com/community/user/businessanalytics/blogs/steven-macko/2019/07/11/telco-customer-churn-1113
