---
title: "Group 7 Project 1: EDA"
author: "Group 7"
date: "3/24/2022"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
options(scientific=T, digits = 3) 
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times


#install packages
#install.packages("DT")
#install.packages("tidyverse")
#install.packages("ezids")
#install.packages("knitr")
#install.packages("kableExtra")
#install.packages("xtable")

#load packages
library(DT)
library(tidyverse)
library(ezids)
library(knitr)
library(kableExtra)
library(xtable)
library(cowplot)
library(ggcorrplot)


```

# 1. Intro

### 1.1 Data Set Overview 

The Data Set is a fictional data set provided by IBM. It provides the demographic information on the customers of a telecommunications company, data on the services each customer used, charges paid (etc.), as well as whether the customer "churned" or not -- meaning, whether the customer left the company. It also includes data on the reason(s) each customer who did leave gave for leaving. 

### 1.2 SMART Questions

Ultimately, the goal is to produce a model for predicting customer churn. This initial project will focus on a basic EDA of the data set in preparation of building such a model. Therefore I want to achieve 2 things:

1. Which of the variables in the data set show a statistical difference between the churn and non-churn groupings?  
-This will require doing 2-sample t-tests for continuous variables, and Chi Squared tests for categorical ones.  
-If a variables passes its tests (i.e. p-value <= 0.05), that would suggest that it had a measurable effect on the decision to churn and can be used for modeling.

2. Of the variables that show a statistical difference between the churn and non-churn groupings, which are strong predictors of churn?
-This will require creating model, for example, a logistical regression, and examining the adjusted r^2 value.  
-This question will be answered in Project 2

3. Of the variables that are strong predictors of churn, which are actually providing useful information?
-This will require critically thinking about what the variable is actually telling us.  
-For example, it is likely that satisfaction score is more or less a corollary of churn and non-churn. Of course customers who churn are unsatisfied.  
-Therefore, even though some variable might be strong predictors, they may not be actually helping us understand churn vs. non-churn.  
-This question will be answered in Project 2

### 1.3 Report Organization and Analysis Steps

1. Prepare and describe data in data set  
2. Divide into churn and non-churn groupings  
3. Graph variables according to churn groupings to see if any obvious visual differences in the variables between churn groups  
4. Conduct appropriate test for each variable to determine if there is a statistical difference between churn groups for each variable 
5. Comment on results and propose next steps

# 2. Description of Data

### 2.1 Load data  and join data sets

*Load Data Sets...*

```{r}
#Main data set
churn_df <- data.frame(read_csv('Data Files/Telco_customer_churn.csv', col_types=cols()))

#additional data sets
churn_status_df <- data.frame(read_csv('Data Files/Telco_customer_churn_demographics.csv', col_types = cols()))
churn_services_df <- data.frame(read_csv('Data Files/Telco_customer_churn_services.csv', col_types = cols()))

```
data set `churn_df` loaded...  
data set `churn_status_df` loaded...  
data set `churn_services_df` loaded...

*Join Data Sets...*
```{r}

status_join <- churn_status_df %>%
  select(c(Customer.ID))
#status_join

#Filer columns to join from services)df
service_join <- churn_services_df %>%
  select(c(Customer.ID, Avg.Monthly.Long.Distance.Charges, Avg.Monthly.GB.Download, Unlimited.Data, Total.Extra.Data.Charges))
#service_join

#Join status with churn
joined_churn_df <- merge(churn_df, status_join, by.x="CustomerID", by.y="Customer.ID")
#head(joined_churn_df)

#Join service with churn
joined_churn_df <- merge(joined_churn_df, service_join, by.x="CustomerID", by.y="Customer.ID")
#head(joined_churn_df)

head(joined_churn_df)

```
*data sets joined...*

### 2.2 Summary of data set structure
```{r results='markup', comment=NA}
#Note -- comment=NA removes the ## normally printed to output before each line

#drop unwanted columns from final df
filtered_joined_churn_df <- joined_churn_df %>%
  select(-c(1:10, 16, 20, 22:23, 28, 30:32))

#display structure of unfiltered df -- so can discuss what was added and dropped
str(joined_churn_df)
```

**Selecting Variables**

The initial churn data frame `(Telco_customer_churn.csv)` contained [`r ncol(churn_df)`]{.underline} variables. Another [`r ncol(service_join)-1`]{.underline} variables were added from the `Telco_customer_churn_services.csv` and  [`r ncol(status_join)-1`]{.underline} variables from the `Telco_customer_churn_status.csv` data set. Conversely, [18]{.underline} variables were dropped/not examined from the original `Telco_customer_churn.csv` data frame. The specific variables added and dropped, and the reason for doing so, follow below:
<br>
<br>

**Added Variables**

*From `Telco_customer_churn_status.csv`*

* ``r names(status_join)[2]``
* ``r names(status_join)[3]``

``r names(status_join)[2]`` was added because it matched reasons customer's gave for choosing to churn. Additionally the variable ``r names(status_join)[3]`` was added to allow for more nuanced examination of the reasons customers gave for choosing to churn. Beyond selecting more variables to use in the investigation, knowing the specific reasons customers gave for leaving the company will be useful to compare against the results of the statistical analysis, in order to better understand and contextualize those results.

*From `Telco_customer_churn_services.csv`*

* ``r names(service_join)[2]``
* ``r names(service_join)[3]``
* ``r names(service_join)[4]``
* ``r names(service_join)[5]``

All of the variables from the `Telco_customer_churn_services.csv` data set were added to the final data frame because they matched the reason customers gave for choosing to churn, as recorded in the `Telco_customer_churn_status.csv` data set.

**Dropped/Not Examined Varaibles**

*From `Telco_customer_churn.csv`*

* ``r names(churn_df)[1]`` [not relevant]
* ``r names(churn_df)[2]`` [useful to ensure uniqueness, but otherwise not relevant]
* ``r names(churn_df)[3]`` [location data not examined]
* ``r names(churn_df)[4]`` [location data not examined]
* ``r names(churn_df)[5]`` [location data not examined]
* ``r names(churn_df)[6]`` [location data not examined]
* ``r names(churn_df)[7]`` [location data not examined]
* ``r names(churn_df)[8]`` [location data not examined]
* ``r names(churn_df)[9]`` [location data not examined]
* ``r names(churn_df)[10]`` [not relevant]
* ``r names(churn_df)[16]`` [assumed corollary with phone service, extra charge from more lines connected with monthly charges]
* ``r names(churn_df)[20]``
* ``r names(churn_df)[22]`` [assumed corollary with GB download and internet speed (type)]
* ``r names(churn_df)[23]`` [assumed corollary with GB download and internet speed (type)]
* ``r names(churn_df)[28]`` [assumed corollary with monthly charges]
* ``r names(churn_df)[30]`` [Same as churn label]
* ``r names(churn_df)[31]`` [this is ultimately what we are trying to determine!]
* ``r names(churn_df)[32]`` [not examining profit, monthly charge more relevant for our purposes]

Most of the variables that were not examined were ignored because they appeared obviously irrelevant, correlated with other variables, or did not seem to provide helpful information with respect to the questions we are trying to answer. Specifically regarding location data,  while examining location data could prove useful, it was decided to use it for follow up analysis. Part of the reason it was not chosen for the primary analysis is this data set does not contain data to dig further into what results on location analysis might mean. We don't have, for example, data on the strength of cell phone networks or availability of internet types by zip code, so it would be hard to usefully connect location data with other data and results.  

<br>
The final data frame, `filtered_joined_churn_df`, consists of [`r ncol(filtered_joined_churn_df)`]{.underline} variables.

**Brief Description of Final Set of Variables**

1.  ``r names(filtered_joined_churn_df)[1]`` [whether customer is a senior citizen]
2.  ``r names(filtered_joined_churn_df)[2]`` [whether customer is single]
3.  ``r names(filtered_joined_churn_df)[3]`` [whether customer has dependents]
4.  ``r names(filtered_joined_churn_df)[4]`` [length of time as customer]
5.  ``r names(filtered_joined_churn_df)[5]`` [whether customer using phone services]
6.  ``r names(filtered_joined_churn_df)[6]`` [type of internet service used -- affects speed]
7.  ``r names(filtered_joined_churn_df)[7]`` [whether customer uses online security service]
8.  ``r names(filtered_joined_churn_df)[8]`` [whether customer uses online backup service]
9.  ``r names(filtered_joined_churn_df)[9]`` [whether customer enrolled in tech support plan]
10. ``r names(filtered_joined_churn_df)[10]`` [type of contract (month-to-monthly, year, etc.)]
11. ``r names(filtered_joined_churn_df)[11]`` [whether customer uses paperless billing]
12. ``r names(filtered_joined_churn_df)[12]`` [payment method that customer uses]
13. ``r names(filtered_joined_churn_df)[13]`` [average total monthly Charge for customer]
14. ``r names(filtered_joined_churn_df)[14]`` [whether a customer did or did not churn]
15. ``r names(filtered_joined_churn_df)[15]`` [specific reason customer gave for churning]
16. ``r names(filtered_joined_churn_df)[16]`` [rating customer's give for their satisfaction as a customer]
17. ``r names(filtered_joined_churn_df)[17]`` [for customer's who churned, category of reason for leaving]
18. ``r names(filtered_joined_churn_df)[18]`` [average long-distance charges in a month]
19. ``r names(filtered_joined_churn_df)[19]`` [avg amount of data downloaded in a month]
20. ``r names(filtered_joined_churn_df)[20]`` [whether customer has unlimited data plan]
21. ``r names(filtered_joined_churn_df)[21]`` [lifetime extra data charges customer has accrued]


### 2.3 Display joined and filted data set
```{r results='markup'}
# filtered_joined_churn_df <- joined_churn_df %>%
#   select(-c(1:10, 16, 20, 22:23, 28, 30:33))

#display final filtered df  
DT::datatable(filtered_joined_churn_df, 
         extensions = c('FixedColumns',"FixedHeader"),
          options = list(scrollX = TRUE, 
                         paging=TRUE,
                         fixedHeader=TRUE))
#xkabledplyhead(joined_churn_df, title = "Churn Data Set")
```

# 3. Examine Data

### 3.1 Loading functions to use for analysis

```{r}
#Create updated xkable_summary function to delete unwanted text
xkablesum_updated <- function (df, title = "Table: Statistics summary.", digits = 4, 
          pos = "left", bso = "striped") 
{
    s = summary(df)
    
    # Including RE NA's strip
    # Needed to add a dim check because including NA strip when no NA row made a new 7th row with text Min.: 
    if (dim(s)[1] == 6) {
      strip_vector = c("Min.\\s*:\\s*", "1st Qu.\\s*:\\s*", "Median\\s*:\\s*", "Mean\\s*:\\s*", 
                       "3rd Qu.\\s*:\\s*", "Max.\\s*:\\s*")
    }
    # NA's strip RE added here
    else if (dim(s)[1] == 7) {
      strip_vector = c("Min.\\s*:\\s*", "1st Qu.\\s*:\\s*", "Median\\s*:\\s*", "Mean\\s*:\\s*", 
                       "3rd Qu.\\s*:\\s*", "Max.\\s*:\\s*", "NA's\\s*:\\s*")
    }
    
    # Made s = apply() -- without, didn't apply changes to table
    s <- apply(s, 2, function(x) stringr::str_remove_all(x, strip_vector))
    
    # Made s = apply()
    s <- apply(s, 2, function(x) stringr::str_trim(x, "right"))
    colnames(s) <- stringr::str_trim(colnames(s))
    
    if (dim(s)[1] == 6) {
        rownames(s) <- c("Min", "Q1", "Median", 
                         "Mean", "Q3", "Max")
    }
    else if (dim(s)[1] == 7) {
        rownames(s) <- c("Min", "Q1", "Median", 
                         "Mean", "Q3", "Max", "NA")
    }
    xkabledply(s, title = title, digits = digits, pos = pos, 
               bso = bso)
}

#Better looking version of 2-sample t-test results than what the object itself displays
ttest2sample_info <- function(test) {
  if (test[["p.value"]] <= 0.05) {
  result = 'Reject the Null Hypothesis'
  } else {
    result = 'Do not reject the Null Hypothesis'
  }
  
  cat(c('\t', test$method, '\n\n',
      'Data:                       ', '|   ', test$data.name, '\n',
      'Null Hypothesis:            ', '|   true difference in means = ', test$null.value[1], '\n',
      'Alternative Hypothesis:     ', '|   true difference in means != ', test$null.value[1], '\n',
      'Confidence Level:           ', '|   ', attributes(test$conf.int)$conf.level, '\n',
      'Confidence Interval:        ', '|   [', round(test$conf.int[1], 2), ', ', round(test$conf.int[2], 2)), ']\n',
      'Sample Estimates of Mean:   ', '|   [X = ', test$estimate[1], ', Y = ', test$estimate[2], ']\n',
      'Test Values:                ', '|   [t = ', test$statistic,
                                   ', |   df = ', test$parameter[1],
                                   ', |   p-value = ', test[["p.value"]], ']\n',
      'Result:                     ', '|   ', result, sep='') 
}

```

*functions loaded...*

### 3.2 Summary Tables of Continuous variables, Divided by Churn Groupings

```{r results='markup'}
#Display summary statistics of continuous variables for entire df


```

```{r}
options(repr.plot.width = 6, repr.plot.height = 4)
joined_churn_df %>% 
group_by(Churn.Label) %>% 
summarise(Count = n())%>% 
mutate(percent = prop.table(Count)*100)%>%
ggplot(aes(reorder(Churn.Label, -percent), percent), fill = Churn)+
geom_col(fill = c("#FC4E07", "#E7B800"))+
geom_text(aes(label = sprintf("%.2f%%", percent)), hjust = 0.01,vjust = -0.5, size =3)+ 
theme_bw()+  
xlab("Churn") + 
ylab("Percent")+
ggtitle("Churn Percent")
```
* around 26% of ustomer left the platform within the last month.

```{r results='markup'}


options(repr.plot.width = 12, repr.plot.height = 8)
plot_grid(ggplot(joined_churn_df, aes(x=Dependents,fill=Churn.Label))+ geom_bar(),ggplot(joined_churn_df, aes(x=Avg.Monthly.GB.Download,fill=Churn.Label))+ geom_bar(position = 'fill')+theme_bw(),
          ggplot(joined_churn_df, aes(x=Unlimited.Data,fill=Churn.Label))+ geom_bar(position = 'fill'),
          ggplot(joined_churn_df, aes(x=Total.Extra.Data.Charges,fill=Churn.Label))+ geom_bar(position = 'fill'),
          ggplot(joined_churn_df, aes(x=Phone.Service,fill=Churn.Label))+ geom_bar(position = 'fill'),
          ggplot(joined_churn_df, aes(x=Internet.Service,fill=Churn.Label))+ geom_bar(position = 'fill')+theme_bw()+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          align = "h")

```

*Customers with Phone service and Dependents have lower churn rate as compared to those who don't have partners & Dependents.
*The percent of churn is higher in case of unlimited data and internet services.


```{r}
options(repr.plot.width = 12, repr.plot.height = 8)
plot_grid(ggplot(joined_churn_df, aes(x=Internet.Service,fill=Churn.Label))+ geom_bar(position = 'fill')+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)), 
          ggplot(joined_churn_df, aes(x=Dependents,fill=Churn.Label))+ geom_bar(position = 'fill')+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          ggplot(joined_churn_df, aes(x=Avg.Monthly.GB.Download,fill=Churn.Label))+ geom_bar(position = 'fill')+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          ggplot(joined_churn_df, aes(x=Unlimited.Data,fill=Churn.Label))+ geom_bar(position = 'fill')+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          ggplot(joined_churn_df, aes(x=Total.Extra.Data.Charges,fill=Churn.Label))+ geom_bar(position = 'fill')+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          ggplot(joined_churn_df, aes(x=Phone.Service,fill=Churn.Label))+ geom_bar(position = 'fill')+theme_bw()+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          align = "h")

```



```{r}
options(repr.plot.width =6, repr.plot.height = 2)
ggplot(joined_churn_df, aes(y= Avg.Monthly.GB.Download, x = "", fill = Churn.Label)) + 
geom_boxplot()+ 
theme_bw()+
xlab(" ")
```

*Customers who left has average gb downloaded more than 10 gb

```{r}
options(repr.plot.width =6, repr.plot.height = 2)
ggplot(joined_churn_df, aes(y= Total.Extra.Data.Charges, x = "", fill = Churn.Label)) + 
geom_boxplot()+ 
theme_bw()+
xlab(" ")
```

```{r}
options(repr.plot.width =6, repr.plot.height = 4)
telco_cor <- round(cor(joined_churn_df[,c("Avg.Monthly.GB.Download", "Total.Extra.Data.Charges")]), 1)

ggcorrplot(telco_cor,  title = "Correlation")+theme(plot.title = element_text(hjust = 0.5))

```
```{r}
cor(telco_cor)

```
*Possitive corr
```{r}
joined_churn_df$Senior.Citizen <- as.factor(joined_churn_df$Dependents)#a
joined_churn_df$Partner <- as.factor(joined_churn_df$Avg.Monthly.GB.Download)
joined_churn_df$Dependents <- as.factor(joined_churn_df$Unlimited.Data)#a
joined_churn_df$Tech.Support <- as.factor(joined_churn_df$Total.Extra.Data.Charges)
joined_churn_df$Internet.Service <- as.factor(joined_churn_df$Phone.Service)#a
joined_churn_df$Online.Security <- as.factor(joined_churn_df$Internet.Service)

```

```{r}
summary(joined_churn_df)
```




# test of independent ( more than three levels)
# Tech.support
```{r ,results='markup'}
x <- table(joined_churn_df$Churn.Label,joined_churn_df$Dependents )
```

```{r,results='hold'}
chitest <- chisq.test(x)
chitest
pvalue<-chitest$p.value
pvalue
#p-value <2e-16, it is not independent. The two variables relate to each other.

```


```{r ,results='markup'}
x <- table(joined_churn_df$Churn.Label,joined_churn_df$Unlimited.Data )
```

```{r,results='hold'}
chitest <- chisq.test(x)
chitest
pvalue<-chitest$p.value
pvalue
#p-value <2e-16, it is not independent. The two variables relate to each other.

```


```{r ,results='markup'}
x <- table(joined_churn_df$Churn.Label,joined_churn_df$Phone.Service )
```

```{r,results='hold'}
chitest <- chisq.test(x)
chitest
pvalue<-chitest$p.value
pvalue
#p-value <2e-16, it is not independent. The two variables relate to each other.

```


```{r ,results='markup'}
x <- table(joined_churn_df$Churn.Label,joined_churn_df$Internet.Service )
```

```{r,results='hold'}
chitest <- chisq.test(x)
chitest
pvalue<-chitest$p.value
pvalue
#p-value <2e-16, it is not independent. The two variables relate to each other.

```

