---
title: "DS_project1"
author: "Jane"
date: "3/22/2022"
output: html_document
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r}
## Importing packages
library(tidyverse) 
library(MASS)
library(car)
library(e1071)
library(caret)
library(cowplot)
library(caTools)
library(pROC)
library(ggcorrplot)
```
## project1
##  the behavior of telecom customers who are more likely to leave the platform. I intend to find out the most striking behavior of customers through EDA

```{r}
telco = data.frame(read.csv("Telco_customer_churn.csv", header = TRUE))
head(telco, n=10)
# 33 variables.
```



```{r}
colnames(telco)
```


```{r}

is.null(telco)
```

```{r}
summary(telco)
```

```{r}
library(Hmisc)
describe(telco)

```




```{r}
library(dplyr)
glimpse(telco)
```
no missing value

```{r}
#telco$Senior.Citizen <- as.factor(ifelse(telco$Senior.Citizen==1, 'YES', 'NO'))

```


```{r}
options(repr.plot.width = 6, repr.plot.height = 4)
telco %>% 
group_by(Churn.Label) %>% 
summarise(Count = n())%>% 
mutate(percent = prop.table(Count)*100)%>%
ggplot(aes(reorder(Churn.Label, -percent), percent), fill = Churn)+
geom_col(fill = c("#FC4E07", "#E7B800"))+
geom_text(aes(label = sprintf("%.2f%%", percent)), hjust = 0.01,vjust = -0.5, size =3)+ 
theme_bw()+  
xlab("Churn") + 
ylab("Percent")+
ggtitle("Churn Percent")
```
* around 26% of ustomer left the platform within the last month.

```{r}
library(cowplot)
options(repr.plot.width = 12, repr.plot.height = 8)
plot_grid(ggplot(telco, aes(x=Gender,fill=Churn.Label))+ geom_bar()+ theme1, 
          ggplot(telco, aes(x=Senior.Citizen,fill=Churn.Label))+ geom_bar(position = 'fill')+theme1,
          ggplot(telco, aes(x=Partner,fill=Churn.Label))+ geom_bar(position = 'fill')+theme1,
          ggplot(telco, aes(x=Dependents,fill=Churn.Label))+ geom_bar(position = 'fill')+theme1,
          ggplot(telco, aes(x=Phone.Service,fill=Churn.Label))+ geom_bar(position = 'fill')+theme1,
          ggplot(telco, aes(x=Multiple.Lines,fill=Churn.Label))+ geom_bar(position = 'fill')+theme_bw()+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          align = "h")
```

```{r}
options(repr.plot.width = 12, repr.plot.height = 8)
plot_grid(ggplot(telco, aes(x=Gender,fill=Churn.Label))+ geom_bar(), 
          ggplot(telco, aes(x=Senior.Citizen,fill=Churn.Label))+ geom_bar(position = 'fill')+theme_bw(),
          ggplot(telco, aes(x=Partner,fill=Churn.Label))+ geom_bar(position = 'fill'),
          ggplot(telco, aes(x=Dependents,fill=Churn.Label))+ geom_bar(position = 'fill'),
          ggplot(telco, aes(x=Phone.Service,fill=Churn.Label))+ geom_bar(position = 'fill'),
          ggplot(telco, aes(x=Multiple.Lines,fill=Churn.Label))+ geom_bar(position = 'fill')+theme_bw()+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          align = "h")

```

*Customers with Partners and Dependents have lower churn rate as compared to those who don't have partners & Dependents.
*The percent of churn is higher in case of senior citizens


```{r}
options(repr.plot.width = 12, repr.plot.height = 8)
plot_grid(ggplot(telco, aes(x=Internet.Service,fill=Churn.Label))+ geom_bar(position = 'fill')+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)), 
          ggplot(telco, aes(x=Online.Security,fill=Churn.Label))+ geom_bar(position = 'fill')+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          ggplot(telco, aes(x=Online.Backup,fill=Churn.Label))+ geom_bar(position = 'fill')+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          ggplot(telco, aes(x=Device.Protection,fill=Churn.Label))+ geom_bar(position = 'fill')+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          ggplot(telco, aes(x=Tech.Support,fill=Churn.Label))+ geom_bar(position = 'fill')+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          ggplot(telco, aes(x=Streaming.TV,fill=Churn.Label))+ geom_bar(position = 'fill')+theme_bw()+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          align = "h")

```

* Churn rate is much higher in case of Fiber Optic InternetServices.

*Customers who do not have services like No OnlineSecurity , OnlineBackup and TechSupport have left the platform in the past month.

```{r}
plot_grid(ggplot(telco, aes(x=Streaming.Movies,fill=Churn.Label))+ 
          geom_bar(position = 'fill')+ theme_bw()+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)), 
          ggplot(telco, aes(x=Contract,fill=Churn.Label))+ 
          geom_bar(position = 'fill')+theme_bw()+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          ggplot(telco, aes(x=Paperless.Billing,fill=Churn.Label))+ 
          geom_bar(position = 'fill')+theme_bw()+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          ggplot(telco, aes(x=Payment.Method,fill=Churn.Label))+
          geom_bar(position = 'fill')+theme_bw()+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          align = "h")

```

*A larger percent of Customers with monthly subscription have left when compared to Customers with one or two year contract.
*Churn percent is higher in case of cutsomers having paperless billing option.
*Customers who have ElectronicCheck PaymentMethod tend to leave the platform more when compared to other options.

in categorical variable:
paperless.billing--yes
payment method--electron check
contract---monthly
InternetServices --Fiber Optic
OnlineSecurity ---no
OnlineBackup--no
TechSupport ---no
Partners---no
Dependents---no
senior---yes???????

```{r}
plot_grid(ggplot(telco, aes(x=Churn.Reason,fill=Churn.Label))+ 
          geom_bar(position = 'fill')+ theme_bw()+
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)), align = "h")
```









```{r}
options(repr.plot.width =6, repr.plot.height = 2)
ggplot(telco, aes(y= Tenure.Months, x = "", fill = Churn.Label)) + 
geom_boxplot()+ 
theme_bw()+
xlab(" ")
```
* Tenure: The median tenure for customers who have left is around 10 months.

```{r}
options(repr.plot.width =6, repr.plot.height = 2)
ggplot(telco, aes(y= Monthly.Charges, x = "", fill = Churn.Label)) + 
geom_boxplot()+ 
theme_bw()+
xlab(" ")
```
*Monthly.Charges: Customers who have churned, have high monthly charges. The median is above 75.

```{r}
options(repr.plot.width =6, repr.plot.height = 2)
ggplot(telco, aes(y= Total.Charges, x = "", fill = Churn.Label)) + 
geom_boxplot()+ 
theme_bw()+
xlab(" ")
```

* Total.Charges: The median Total charges of customers who have churned is low.


```{r}
options(repr.plot.width =6, repr.plot.height = 4)
telco_cor <- round(cor(telco[,c("Tenure.Months", "Monthly.Charges", "Total.Charges")]), 1)

ggcorrplot(telco_cor,  title = "Correlation")+theme(plot.title = element_text(hjust = 0.5))

```
* positive corr

paperless.billing--yes
payment method--electron check
contract---monthly
InternetServices --Fiber Optic
OnlineSecurity ---no
OnlineBackup--no
TechSupport ---no
Partners---no
Dependents---no
senior---yes???????
```{r}
glimpse(telco)
```

```{r}
telco$Senior.Citizen <- as.factor(telco$Senior)
telco$Partner <- as.factor(telco$Partner)
telco$Dependents <- as.factor(telco$Dependents)
telco$Tech.Support <- as.factor(telco$Tech)
telco$Internet.Service <- as.factor(telco$Internet)
telco$Online.Security <- as.factor(telco$Online.Security)
telco$Online.Backup <- as.factor(telco$Online.Backup)
telco$Paperless.Billing <- as.factor(telco$Paperless.Billing)
telco$contract <- as.factor(telco$contract)
telco$Payment.Method <- as.factor(telco$Payment.Method)
```
















